---
title: "IDA Hw1"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r}
z1 = data.frame(rnorm(500,0,1), stringsAsFactors=FALSE) # generated three list of data that are normally distributed
z2 = data.frame(rnorm(500,0,1), stringsAsFactors=FALSE)
z3 = data.frame(rnorm(500,0,1), stringsAsFactors=FALSE)
Y1 = data.frame(z1+1, stringsAsFactors = FALSE)
Y2 = data.frame(5+2*z1+z2)
Y2[2*(z1-1)+z3<0] = NA
Y3 = data.frame(5+2*z1+z2) #Y3 is fully observed data of Y2 after imposing missingness
data = data.frame(Y1,Y2)
colnames(data) = c("Y1", "Y2")
data
require(VIM)
marginplot(data[c(1,2)]) # plot of the marginal distribution of Y2 with missing values
data1 = data.frame(Y1,Y3)
colnames(data1) = c("Y1", "Y3")
data1
marginplot(data1[c(1,2)]) # plot of the marginal distribution of Y2 after imposing missingness
```


# Question 3 part 2
```{r setup2, include=TRUE}
set.seed(124)
fit1 = lm(data$Y2 ~ data$Y1, data = data)
stochastic_1 = predict(fit1, newdata = data) + rnorm(nrow(data), 0, sigma(fit1))
data2 = data.frame(Y1,stochastic_1)
colnames(data2) = c("Y1", "Y2")
data2
marginplot(data2[c(1,2)])
```

#Question 3 part 3
```{r}
y1 = data.frame(z1+1, stringsAsFactors = FALSE)
y2 = data.frame(5+2*z1+z2)
y2[2*(5+2*z1+z2-1)+z3<0] = NA
y3 = data.frame(5+2*z1+z2) 
data3 = data.frame(y1,y2)
colnames(data3) = c("Y1", "Y2")
data3

marginplot(data3[c(1,2)]) 

data4 = data.frame(Y1,Y3)
colnames(data4) = c("Y1", "Y3")
data4
marginplot(data4[c(1,2)]) 

```
# Question 3 Part 4
```{r}
set.seed(124)
fit2 = lm(data3$Y2 ~ data3$Y1, data = data3)
stochastic_2 = predict(fit2, newdata = data3) + rnorm(nrow(data3), 0, sigma(fit2))
data5 = data.frame(y1,stochastic_2)
colnames(data5) = c("Y1", "Y2")
data5
marginplot(data5[c(1,2)])
```


#Question 4 Part 1
```{r cars}
patients = get(load('databp.Rdata'))
patients
mean(patients$recovtime,na.rm=TRUE)
cor(patients$recovtime,patients$logdose, use = "complete.obs")
cor(patients$recovtime,patients$bloodp, use = "complete.obs")

```
# Question 4 part 2
```{r}
recov_mis = ifelse(is.na(patients$recovtime), mean(patients$recovtime,na.rm=TRUE), patients$recovtime)
mean(recov_mis)
cor(recov_mis,patients$logdose,use = "complete")
cor(recov_mis,patients$bloodp,use = "complete")
```

# Question 4 part 3
```{r pressure, echo=FALSE}
fit = lm(patients$recovtime ~ patients$logdose + patients$bloodp, data = patients)
summary(fit)
linear = predict(fit, newdata = patients)
mean(linear)
cor(linear,patients$logdose,use = "complete")
cor(linear,patients$bloodp,use = "complete")
```
# Question 4 part 4
```{r}
set.seed(124)
stochastic = predict(fit, newdata = patients) + rnorm(nrow(patients), 0, sigma(fit))
mean(stochastic)
cor(stochastic,patients$logdose,use = "complete")
cor(stochastic,patients$bloodp,use = "complete")

```
# Question 5 part 5
```{r}
# I wasn't quite sure what the question means. Based on my understanding, the mean matching method is to replace the missing values by donors with their observed values being closest to the simulated value of each missing row. 

available = data.frame(patients$recovtime, stringsAsFactors = FALSE)
aa = remove_missing(available, na.rm = TRUE) # First I removed missing values from the recover time column
aa
which(is.na(available)) # Then I located the index of each missing value's location.

# I looked up the values from the observed data and find the closest one to each of the simulated data from regression imputation method.
donor_1 = aa[which.min(abs(c(aa[1:22]) - linear[4]))] 
donor_2 = aa[which.min(abs(c(aa[1:22]) - linear[10]))]
donor_3 = aa[which.min(abs(c(aa[1:22]) - linear[22]))]
# Next I stored the three donor values and the observed data in one vector and resorted them to the correct order.
mean_match = c(donor_1,donor_2,donor_3,aa)
aa2 = append(aa, 4, values=donor_1)
aa3 = append(aa2, 9, values=donor_2)
aa4 = append(aa3, 21, values=donor_3)
aa4
mean(aa4)
cor(aa4,patients$logdose,use = "complete")
cor(aa4,patients$bloodp,use = "complete")

```
# Question 5 part 6

One of the aspects that predicted mean matching is better than stochastic regression imputation is that it generates more accurate mean. One  of the potential problem for this method is it can be extremely inaccurate when there are too many missing values. If there are more missing values that observed, the value of donors may be significantly different than the predicted value from regression imputation.
